# Labor force market in Puerto Rico and U.S. mainland
##### The data I am using in this research is ACS (American Community Surveys) 2016 data, From the data, I found that the labor force participation rate in Puerto Rico(0.58) is far lower than it in the U.S. mainland(0.75).
##### To find out the reason that caused such a huge difference, I took age and disability as two main variables and tried to explore the relationship between these variables and the labor force statue.


# Get the data from Github
```{r}
data0URL <- "https://raw.githubusercontent.com/TianXinyu/FinalProject/master/usa_00003.csv"
data0 <- read.csv(data0URL)
```

# See the structure of the data
```{r}
class(data0)
names(data0)
str(data0)
```

# Clean the dataset, delete all the missing values and only keep the variables you need.
## 1. Make age a numerical variable
```{r}
data0$AGE=as.numeric(data0$AGE)
data0$DATANUM=as.numeric(data0$DATANUM)
str(data0$AGE)
str(data0$DATANUM)
```

## 2. Keep people from 25-65
```{r}
data0=data0[data0$AGE>=25 & data0$AGE<=65, ]
str(data0)
```

## 3. Clean all the missing datas in labor force, 0 is not in lbfc, 1 is in lbfc
```{r}
data0$LABFORCE=as.numeric(data0$LABFORCE)
data0$lbfc[data0$LABFORCE==1] <- 0
data0$lbfc[data0$LABFORCE==2] <- 1
str(data0)
```

## 4. Combine the diseases column together, get a new veriable(column), o is healthy, 1 is disabled
```{r}
Disable <- rep_len(0, length.out=nrow(data0))
data0 <- cbind(data0, Disable)
data0$Disable <- data0$Disable[data0$DIFFPHYS==1 & data0$DIFFCARE==1 & data0$DIFFMOB==1 & data0$DIFFSENS==1 & data0$DIFFEYE==1 & data0$DIFFHEAR==1] <-0
data0$Disable[data0$DIFFPHYS==2 | data0$DIFFCARE==2 | data0$DIFFMOB==2 |  data0$DIFFSENS==2 | data0$DIFFEYE==2 |  data0$DIFFHEAR==2] <-1
str(data0)
```

# Get a new dataframe after cleaning
```{r}
data1 <- data0[,c('DATANUM','AGE','lbfc','Disable')]
str(data1)
```

# Seperate the data into US data and PR data
```{r}
prdata <- data1[data1$DATANUM==2,]
str(prdata)
usdata <- data1[data1$DATANUM==1,]
str(usdata)
```

# Use probiblity table to see the labor force participation rate in PR and US
```{r}
prop.table(table(prdata$lbfc))
prop.table(table(usdata$lbfc))
```

# Explore the relationship between lbfc and age.
## 1. Use boxplot to show the average age of both lbfc statues in PR
```{r}
boxplot(prdata$AGE~prdata$lbfc)
```

## 2. Sepret age into Y (young25-45) and O(old,45-65)
```{r}
prY <- prdata[prdata$AGE<=45,]
prO <- prdata[prdata$AGE>45,]
usY <- usdata[usdata$AGE<=45,]
usO <- usdata[usdata$AGE>45,]
```

## 3. Compare the difference between young people
```{r}
prop.table(table(prY$lbfc))
prop.table(table(usY$lbfc))
```

## 4. Compare the difference between old people
```{r}
prop.table(table(prO$lbfc))
prop.table(table(usO$lbfc))
```

# Explore the relationship between lbfc and disability
## 1. Use probiblity table to see the disability rate in PR and US
```{r}
prop.table(table(prdata$Disable))
prop.table(table(usdata$Disable))
```

## 2. Use boxplot to show the average age of both health statues in PR
```{r}
boxplot(prdata$AGE~prdata$Disable)
```

## 3. For people who are disabled, how many of them are in labor force
```{r}
prdis <- prdata[prdata$Disable==1,]
usdis <- usdata[usdata$Disable==1,]
prop.table(table(prdis$lbfc))
prop.table(table(usdis$lbfc))
```

# Use regression analysis to confirm the relationship between age and disability.
## 1. PR
```{r}
prdata$Disable=as.factor(prdata$Disable)
mod=glm(lbfc~.,data=prdata,family = 'binomial')
summary(mod)
```

## 2. US
```{r}
usdata$Disable=as.factor(usdata$Disable)
mod1=glm(lbfc~.,data=usdata,family = 'binomial')
summary(mod1)
```

##### At last, I want to say "thank you" to professor Magallanes, Steven and Isabella, I would't be able to finish this simple project without your generous help.